---
title: "Reproducible Research: Peer Assessment 1"
author: "Jorge Márquez"
date: "April, 2017"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Load data into data frame `activity`. 
```{r}
setwd('C:/Users/Toshiba/Documents/R/Coursera/Activity')
activity <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
Calculate sum of steps per day, create histogram
```{r} 
steps_per_day <- aggregate(steps ~ date, activity, sum)
hist(steps_per_day$steps, 
     main = "Total Steps Taken per Day", 
     col="blue", 
     xlab="Number of Steps")
```
Calculate mean and median.
```{r} 
totalmean <- mean(steps_per_day$steps)
totalmedian <- median(steps_per_day$steps)
```

* The mean is
```{r}
totalmean
```
* The median is
```{r} 
totalmedian
```


## What is the average daily activity pattern?
Calculate average steps per day. 
Plot the average steps in an interval of time. 
```{r}
steps_by_interval <- aggregate(steps ~ interval, activity, mean)
plot(steps_by_interval$interval,
     steps_by_interval$steps, 
     type="l", 
     xlab="5-Minute Interval", 
     ylab="Average Number of Steps",
     main="Average Number of Steps per Day by Interval")
```

Find interval with max steps interval.
```{r}
max_steps_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
```
* The maximum number of steps is
```{r}
max_steps_interval 
```

## Imputing missing values
Missing data needed to be imputed. 
```{r}
incomplete <- sum(!complete.cases(activity))
imputed_data <- transform(activity, 
                          steps = ifelse(is.na(activity$steps), steps_by_interval$steps[match(activity$interval, steps_by_interval$interval)], activity$steps))
```

Zeroes were imputed for 10-01-2012.
```{r}
imputed_data[as.character(imputed_data$date) == "2012-10-01", 1] <- 0
```

Recount total steps per day and create histogram. 
```{r}
steps_per_day_new <- aggregate(steps ~ date, imputed_data, sum)
hist(steps_per_day_new$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")

hist(steps_per_day$steps, main = paste("Total Steps Each Day"), col="red", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "red"), lwd=10)
```

Calculate new mean and median for imputed data. 
```{r}
newmean <- mean(steps_per_day_new$steps)
newmedian <- median(steps_per_day_new$steps)
```

* The new mean for imputed data is
```{r}
newmean
```
* The new median for imputed data is
```{r}
newmedian 
```

Calculate difference between imputed and non-imputed data.
```{r}
meandiff <- newmean - totalmean
mediandiff <- newmedian - totalmedian
```

* The difference between the non-imputed mean and imputed mean is
```{r}
meandiff
```
* The difference between the non-imputed mean and imputed median is
```{r}
mediandiff
```

Calculate total difference.
```{r}
totaldiff <- sum(steps_per_day_new$steps) - sum(steps_per_day$steps)
```

* The difference between total number of steps between imputed and non-imputed data is
```{r}
totaldiff
```

## Are there differences in activity patterns between weekdays and weekends?
Created a plot to compare and contrast number of steps between the week and weekend.
```{r}
weekends <- c("sábado", "domingo")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekends), "Weekend", "Weekday"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)

library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Number of Steps",layout=c(1,2), type="l")

```